---
title: "Plasmid_and_cassettes_insertions_excluding_low_relAbund"
output: pdf_document
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes: 
  \usepackage{longtable}
---

AVVenger was run on Aradhana's sequences (210426\_M03249\_0176\_000000000-G7P6T) using a database that includes the dog genome, the plasmid backbone, and the Heavy and light chain cassettes (dog\_plus\_plasmid_parts.2bit). Options used were ```sequence.qualTrim.code: '+' , alignment.removeVectorReadPairs: FALSE```. All sites with Sonic abundance less than 1 were excluded from the analysis.

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
suppressMessages(library("tidyverse"))
library(kableExtra)

get_insertion_type <- function(x) {
  xx<-as.character(x)
  x <- ifelse(startsWith(xx, 'chr'),'dog',
    ifelse(startsWith(xx, 'LC_plasmid_backbone'),"plasmid_BB",
      ifelse(startsWith(xx, 'LightChain'),'LightChain','HeavyChain')
    )
  )
  return(x)
}

all_sites<- readRDS('dog_plus_plasmid/out1/sites.rds')

tmp_sites <- all_sites %>%
  mutate(insertion_type=
           as.factor(get_insertion_type(seqnames)),
         readID.list=NULL,
         ltrRepSeq=NULL,
         ltrRepSeq2=NULL) 


all_sites_clean <- tmp_sites %>%
  filter(estAbund>1) %>%
  group_by(sample,replicate) %>% 
  mutate(relAbund=estAbund/sum(estAbund)) %>% 
  ungroup() %>% 
  arrange(sample,replicate)


all_sites_summ <- all_sites_clean %>% group_by(sample,replicate) %>% 
  summarise(dog_relAbund=sum(relAbund*ifelse(insertion_type=="dog",1,0)),
            plasmid_BB_relAbund=sum(relAbund*ifelse(insertion_type=="plasmid_BB",1,0)),
            LC_relAbund=sum(relAbund*ifelse(insertion_type=="LightChain",1,0)),
            HC_relAbund=sum(relAbund*ifelse(insertion_type=="HeavyChain",1,0))
            ) %>%
  mutate_if(is.numeric, ~replace(., is.na(.), 0)) %>%
  ungroup()

long_all_sites_summ <- all_sites_summ %>%
  gather(insertion_type,relAbund, plasmid_BB_relAbund,
         dog_relAbund,LC_relAbund,HC_relAbund)
```

# Insertion sites detected

No insertion sites where detected in the control samples (as expected). Only insertion sites to the plasmid and corresponding cassette were detected in the plasmid samples (also as expected). Dog, plasmid backbone and insertion to both cassettes were detected in one of the dog samples. 

```{r sample_table, results='asis'}
library(xtable)
all_samples <-c("GTSP4175","GTSP4176","GTSP4181","GTSP4182",
                "UninfectedControl-210422-MN","NoTemplateControl-210422-MN")
sample_type <-c("plasmid","plasmid","dog","dog","control","control")

count_ins_dog <- sapply(all_samples,function(x){sum(
  (all_sites_clean$sample==x) & (all_sites_clean$insertion_type=='dog'))})
count_ins_plasmid_BB <- sapply(all_samples,function(x){sum(
  (all_sites_clean$sample==x) & (all_sites_clean$insertion_type=='plasmid_BB'))})
count_ins_LC <- sapply(all_samples,function(x){sum(
  (all_sites_clean$sample==x) & (all_sites_clean$insertion_type=='LightChain'))})
count_ins_HC <- sapply(all_samples,function(x){sum(
  (all_sites_clean$sample==x) & (all_sites_clean$insertion_type=='HeavyChain'))})

sample_desc <- as.data.frame(t(rbind(sample_type,count_ins_dog,
                                     count_ins_plasmid_BB,count_ins_LC,count_ins_HC)))
colnames(sample_desc) <- c("Sample type","dog","plasmid BB","LC","HC")
pp_table <- xtable(sample_desc,caption = c('Counts of insertion sites'))
align(pp_table) <- "|l|ccccc|"
print(pp_table,comment=F)
```


# Relative abundance

Relative abundance of dog insertion sites in the dog samples ranges from 57\% to 68\%

```{r table,results='asis'}

print_table <- xtable(all_sites_summ)
align(print_table) <- "|l|lrrrrr|"
print(print_table, comment=F,include.rownames = FALSE,
      hline.after = c(-1,0,2,5,9),floating=FALSE)

```

# Boxplot

```{r boxplot}
ggplot(long_all_sites_summ, aes(x=sample, y=relAbund,fill=insertion_type)) + 
  geom_boxplot()

```


# Distances (mess-upness)

For each sample, we have four measurements (relative abundance of dog, HC transgene, LC transgene and Plasmid backbone) that add to 1. we can generate a 4D vector with those numbers e.g. (0.62,0.02,0.16,0.20) for sample STSP4181 replicate 4. We also have a expected (or ideal) vector e.g. (1,0,0,0) (only dog genome insertion) for the same sample. Distances between the two vectors gives us an estimate of the mess-upness

```{r m_table,results='asis'}

dog_mesup <- all_sites_summ %>% filter(sample=="GTSP4181") %>% 
  transmute(sample=sample, replicate=replicate,
    mesupness= sqrt((dog_relAbund-1)^2+plasmid_BB_relAbund^2+LC_relAbund^2+HC_relAbund^2))

dog_mesup %>%
  kable( "latex", booktabs = TRUE, longtable = TRUE, caption = "Dog sample messupness") %>%
  kable_styling(latex_options = c("hold_position", "repeat_header"))

```

# Position of instetions into the plasmid backbone

Insertion events can be observed across the length of the plasmid. Insertions located at the start and end might be the expected construct.

```{r l_table,results='asis'}
names(sample_type) <- all_samples
sites_l_table <- all_sites %>% filter(!startsWith(seqnames, 'chr') & estAbund>1) %>% 
  #all_sites[!startsWith(all_sites$seqnames, 'chr'),] 
  transmute(sample=sample,
            sample_type=sample_type[sample],
            replicate=replicate,
            insertion_type=as.factor(get_insertion_type(seqnames)),
            position=position,
            strand=strand,
            estAbund=estAbund) %>%
arrange(position) %>% mutate(position=paste0(position,strand),strand=NULL)

sites_l_table[sites_l_table$insertion_type=="plasmid_BB",] %>% mutate(insertion_type = NULL) %>%
kable( "latex", booktabs = TRUE, longtable = TRUE, caption = "Plasmid backbone instertions") %>%
  kable_styling(latex_options = c("hold_position", "repeat_header"))

```


# Position of insertion into the Heavy chain cassette

```{r l_table_HC,results='asis'}
sites_l_table %>% filter(insertion_type=="HeavyChain") %>% mutate(insertion_type = NULL) %>%
kable( "latex", booktabs = TRUE, longtable = TRUE, caption = "Heavy chain cassette instertions") %>%
  kable_styling(latex_options = c("hold_position", "repeat_header"))
```

\newpage

# Position of insertion into the Light chain cassette

```{r l_table_LC,results='asis'}
sites_l_table %>% filter(insertion_type=="LightChain") %>% mutate(insertion_type = NULL) %>%
kable( "latex", booktabs = TRUE, longtable = TRUE, caption = "Light chain cassette instertions") %>%
  kable_styling(latex_options = c("hold_position", "repeat_header"))
```

```{r l_dog,results='asis'}

library(gt23)
suppressMessages(library(BSgenome.Cfamiliaris.UCSC.canFam3))

dog_sites_l_table_tmp <- all_sites %>% filter(startsWith(seqnames, 'chr') & estAbund>1)
dog_sites_l_table_tmp$refGenome <- 'mm9'
dog_sites_l_table_tmp$start <- dog_sites_l_table_tmp$position
dog_sites_l_table_tmp$end <- dog_sites_l_table_tmp$position
 
s <- GenomicRanges::makeGRangesFromDataFrame(data.frame(dog_sites_l_table_tmp), keep.extra.columns = TRUE)
suppressMessages(s <- gt23::annotateIntSites(s))
s <- data.frame(s)

  
dog_sites_l_table <- s %>% transmute(sample=sample,
            sample_type=sample_type[sample],
            replicate=replicate,
            insertion_type=as.factor(get_insertion_type(seqnames)),
            position=position,
            strand=strand,
            estAbund=estAbund,
            inFeature=inFeature,
            nearestFeature=nearestFeature,
            nearestOncoFeature=nearestOncoFeature,
            seqnames=seqnames
            ) %>%
            arrange(seqnames,position) %>%
            mutate(position=paste0(seqnames,strand,position),strand=NULL,seqnames=NULL,insertion_type = NULL,sample_type=NULL)
            
  
dog_sites_l_table %>%  kable( "latex", booktabs = TRUE, longtable = TRUE, caption = "Dog instertions") %>%
            kable_styling(latex_options = c("hold_position", "repeat_header"))


```

